Functional pant indicators - wetland, indicator documentation

### 2023-04-17 DRAFT_functional plant indicators_wetland.Rmd
tmap wetland analyses

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(tidyr)
library(dplyr)
library(plyr)
library(stringr)
library(tidyverse)
library(readxl)
library(broom)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(ggplot2)
library(tmap)
library(raster)
library(stars)

@@

```{r}
head(results.wetland[['2-sided']])
```

>>>>

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(tidyr)
library(dplyr)
library(plyr)
library(stringr)
library(tidyverse)
library(readxl)
library(broom)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(ggplot2)
library(tmap)
library(tmaptools)
library(raster)
library(stars)

@@

```{r}
head(results.wetland[['2-sided']])
```
#### Scaled value analyses

```{r}
#### plotting scaled values by main ecosystem type ####
## continuing with 2-sided
res.wetland <- results.wetland[['2-sided']]

# make long version of the scaled value part
res.wetland <-
  res.wetland %>% 
  pivot_longer(
    cols = c("Light1","Light2","Moist1","Moist2","pH1","pH2","Nitrogen1","Nitrogen2"),
    names_to = "fp_ind",
    values_to = "scaled_value",
    values_drop_na = TRUE
  )

# summarizing the indicator scores
res.wetland %>%
  group_by(fp_ind) %>%
  dplyr::summarize(Mean = mean(scaled_value, na.rm=TRUE))

# making the plot
ggplot(res.wetland, aes(x=factor(hovedtype_rute), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Main ecosystem type") + 
  ylab("Scaled indicator value") 
```

We can also show the results as a map, for instance for pH1 (the lower pH indicator), either by directly plotting the data onto the map...
```{r}
#### scaled value maps ####
# keep wide format and add geometry again
res.wetland2 <- results.wetland[['2-sided']]
st_geometry(res.wetland2) <- st_geometry(ANO.wetland)

nor <- readRDS('P:/41201785_okologisk_tilstand_2022_2023/data/rds/norway_outline.RDS')%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))
reg <- st_read("P:/41201785_okologisk_tilstand_2022_2023/data/regioner/regNorway_wgs84 - MERGED.shp")%>%
  st_as_sf() %>%
  st_transform(crs = crs(ANO.geo))
# change region names to something R-friendly
reg$region
reg$region <- c("Northern Norway","Central Norway","Eastern Norway","Western Norway","Southern Norway")

regnor <- st_intersection(reg,nor)

## scaled value maps
# pH1 (lower indicator)
tm_shape(regnor) +
  tm_fill('GID_0', labels="", title="") + tm_borders() +
  tm_shape(res.wetland2) +
  tm_dots('pH1',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=1, legend.show = FALSE) + # 
  tm_layout(main.title = "Moisture index (lower), wetland",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 5, plot = FALSE),'grey'),
                labels = c("0.4", "...", "...", 
                           "...", "1", "NA"),
                title = "index values")
...or by showing the region-wise means and standard deviations
```{r}
regnor <- regnor %>%
  mutate(Light1.reg.mean = c(mean(res.wetland2$Light1[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$Light1[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$Light1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$Light1[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$Light1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Light2.reg.mean = c(mean(res.wetland2$Light2[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$Light2[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$Light2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$Light2[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$Light2[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Moist1.reg.mean = c(mean(res.wetland2$Moist1[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$Moist1[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$Moist1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$Moist1[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$Moist1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Moist2.reg.mean = c(mean(res.wetland2$Moist2[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$Moist2[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$Moist2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$Moist2[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$Moist2[res.wetland2$region=="Southern Norway"],na.rm=T)),
         pH1.reg.mean = c(mean(res.wetland2$pH1[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$pH1[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$pH1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$pH1[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$pH1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         pH2.reg.mean = c(mean(res.wetland2$pH2[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$pH2[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$pH2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$pH2[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$pH2[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Nitrogen1.reg.mean = c(mean(res.wetland2$Nitrogen1[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen1[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen1[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Nitrogen2.reg.mean = c(mean(res.wetland2$Nitrogen2[res.wetland2$region=="Northern Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen2[res.wetland2$region=="Central Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen2[res.wetland2$region=="Western Norway"],na.rm=T),
                             mean(res.wetland2$Nitrogen2[res.wetland2$region=="Southern Norway"],na.rm=T)),

         Light1.reg.sd = c(sd(res.wetland2$Light1[res.wetland2$region=="Northern Norway"],na.rm=T),
                             sd(res.wetland2$Light1[res.wetland2$region=="Central Norway"],na.rm=T),
                             sd(res.wetland2$Light1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             sd(res.wetland2$Light1[res.wetland2$region=="Western Norway"],na.rm=T),
                             sd(res.wetland2$Light1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Light2.reg.sd = c(sd(res.wetland2$Light2[res.wetland2$region=="Northern Norway"],na.rm=T),
                             sd(res.wetland2$Light2[res.wetland2$region=="Central Norway"],na.rm=T),
                             sd(res.wetland2$Light2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             sd(res.wetland2$Light2[res.wetland2$region=="Western Norway"],na.rm=T),
                             sd(res.wetland2$Light2[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Moist1.reg.sd = c(sd(res.wetland2$Moist1[res.wetland2$region=="Northern Norway"],na.rm=T),
                             sd(res.wetland2$Moist1[res.wetland2$region=="Central Norway"],na.rm=T),
                             sd(res.wetland2$Moist1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             sd(res.wetland2$Moist1[res.wetland2$region=="Western Norway"],na.rm=T),
                             sd(res.wetland2$Moist1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Moist2.reg.sd = c(sd(res.wetland2$Moist2[res.wetland2$region=="Northern Norway"],na.rm=T),
                             sd(res.wetland2$Moist2[res.wetland2$region=="Central Norway"],na.rm=T),
                             sd(res.wetland2$Moist2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                             sd(res.wetland2$Moist2[res.wetland2$region=="Western Norway"],na.rm=T),
                             sd(res.wetland2$Moist2[res.wetland2$region=="Southern Norway"],na.rm=T)),
         pH1.reg.sd = c(sd(res.wetland2$pH1[res.wetland2$region=="Northern Norway"],na.rm=T),
                          sd(res.wetland2$pH1[res.wetland2$region=="Central Norway"],na.rm=T),
                          sd(res.wetland2$pH1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                          sd(res.wetland2$pH1[res.wetland2$region=="Western Norway"],na.rm=T),
                          sd(res.wetland2$pH1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         pH2.reg.sd = c(sd(res.wetland2$pH2[res.wetland2$region=="Northern Norway"],na.rm=T),
                          sd(res.wetland2$pH2[res.wetland2$region=="Central Norway"],na.rm=T),
                          sd(res.wetland2$pH2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                          sd(res.wetland2$pH2[res.wetland2$region=="Western Norway"],na.rm=T),
                          sd(res.wetland2$pH2[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Nitrogen1.reg.sd = c(sd(res.wetland2$Nitrogen1[res.wetland2$region=="Northern Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen1[res.wetland2$region=="Central Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen1[res.wetland2$region=="Eastern Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen1[res.wetland2$region=="Western Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen1[res.wetland2$region=="Southern Norway"],na.rm=T)),
         Nitrogen2.reg.sd = c(sd(res.wetland2$Nitrogen2[res.wetland2$region=="Northern Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen2[res.wetland2$region=="Central Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen2[res.wetland2$region=="Eastern Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen2[res.wetland2$region=="Western Norway"],na.rm=T),
                                sd(res.wetland2$Nitrogen2[res.wetland2$region=="Southern Norway"],na.rm=T)),

         Light1.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$Light1),]),
                             nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$Light1),]),
                             nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$Light1),]),
                             nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$Light1),]),
                             nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$Light1),])),
         Light2.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$Light2),]),
                             nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$Light2),]),
                             nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$Light2),]),
                             nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$Light2),]),
                             nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$Light2),])),
         Moist1.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$Moist1),]),
                             nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$Moist1),]),
                             nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$Moist1),]),
                             nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$Moist1),]),
                             nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$Moist1),])),
         Moist2.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$Moist2),]),
                             nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$Moist2),]),
                             nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$Moist2),]),
                             nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$Moist2),]),
                             nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$Moist2),])),
         pH1.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$pH1),]),
                          nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$pH1),]),
                          nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$pH1),]),
                          nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$pH1),]),
                          nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$pH1),])),
         pH2.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$pH2),]),
                          nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$pH2),]),
                          nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$pH2),]),
                          nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$pH2),]),
                          nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$pH2),])),
         Nitrogen1.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$Nitrogen1),]),
                                nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$Nitrogen1),]),
                                nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$Nitrogen1),]),
                                nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$Nitrogen1),]),
                                nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$Nitrogen1),])),
         Nitrogen2.reg.n = c(nrow(res.wetland2[res.wetland2$region=="Northern Norway" & !is.na(res.wetland2$Nitrogen2),]),
                                nrow(res.wetland2[res.wetland2$region=="Central Norway" & !is.na(res.wetland2$Nitrogen2),]),
                                nrow(res.wetland2[res.wetland2$region=="Eastern Norway" & !is.na(res.wetland2$Nitrogen2),]),
                                nrow(res.wetland2[res.wetland2$region=="Western Norway" & !is.na(res.wetland2$Nitrogen2),]),
                                nrow(res.wetland2[res.wetland2$region=="Southern Norway" & !is.na(res.wetland2$Nitrogen2),]))
         )

## scaled value maps
# pH1 (lower indicator)
tm_shape(regnor) +
  tm_polygons(col="pH1.reg.mean", title="pH (lower), mean", style="quantile", palette=rev(get_brewer_pal(palette="OrRd", n=5, plot=FALSE)))

tm_shape(regnor) +
  tm_polygons(col="pH1.reg.sd", title="pH (lower), sd", style="quantile", palette=get_brewer_pal(palette="OrRd", n=5, plot=FALSE))
